\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[hidelinks]{hyperref}
\usepackage{doi}

\input{glyphtounicode}
\pdfgentounicode=1

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{remark}
\newtheorem{remark}{Remark}

\newcommand{\JPi}{J_{\Pi}}
\newcommand{\Jthree}{J_{3}}
\newcommand{\Cgen}{\mathbb{C}^{3}_{\mathrm{gen}}}
\newcommand{\Epi}{E_{\Pi}}
\newcommand{\Sbundle}{\mathcal{S}_{\Pi}}
\newcommand{\SL}{\mathcal{S}_{L,\Pi}}
\newcommand{\SR}{\mathcal{S}_{R,\Pi}}
\newcommand{\LY}{L_{Y}}
\newcommand{\Ypi}{Y_{\Pi}}
\newcommand{\Hpi}{H_{\Pi}}
\newcommand{\Upi}{U_{\Pi}}
\newcommand{\lamY}{\lambda_{Y}}
\newcommand{\Sym}{\mathrm{Sym}}
\newcommand{\SU}{\mathrm{SU}}
\newcommand{\UU}{\mathrm{U}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\rank}{\operatorname{rank}}

\title{
  Projected Yukawa Operator and Chiral Polar Factor
  \\[1ex]
  \large The Squared Observable $\Hpi = \Ypi^{\dagger}\Ypi$ and the Undetermined Chiral Polar Map $\Upi$
}

\author{J. Beau\\
Independent Researcher, France\\
\texttt{jerome.beau@cosmochrony.org}}

\date{2026}

\begin{document}

  \maketitle

  \begin{abstract}
    This note takes the second step of the mass-sector frontier of the fermionic-matter sub-programme of
    Cosmochrony, and draws a sharp negative conclusion at its entry.
    The companion line note fixed the determinant line $\LY = \wedge^{2}(\Sbundle)$ as the Yukawa coupling
    line and the generation-level assignment from $\Epi^{2}|_{\Cgen} = \diag(1, \tfrac12 + u, \tfrac12 -
    u)$~\cite{Beau2026pyl}.
    We ask whether the squared projective residue determines the Yukawa morphism itself, and show that it
    does not.
    The generation-level observable produced by the Schur-residue sector is the positive hermitian square
    $\Hpi := \Ypi^{\dagger}\Ypi$, whose spectrum on $\Cgen$ is fixed up to the overall Yukawa norm by
    $\Hpi|_{\Cgen} = \lamY^{2}\,\diag(1, \tfrac12 + u, \tfrac12 - u)$.
    The morphism $\Ypi$ is then determined only up to a chiral polar factor, $\Ypi = \Upi\,\Hpi^{1/2}$, with
    $\Upi$ a unitary chiral map on the support of $\Hpi$.
    Hence the squared residue fixes the squared Yukawa levels but not the morphism: $\Upi$ carries the chiral
    orientation, the relative generation phases, and the mixing, while the positive scale $\lamY$ is a
    separate normalisation of $\Hpi$ --- none of it visible to the dimensionless level operator
    $\Epi^{2}|_{\Cgen}$.
    Front~3b is therefore closed as a squared-Yukawa-levels result, not a mixing result.
    A further analysis characterises $\Upi$ not as a canonical matrix but as a rephasing class, the double
    quotient $[\Upi] \in \UU(1)^{3}_{R} \backslash \UU(3) / \UU(1)^{3}_{L}$, whose invariants are the moduli
    $|(\Upi)_{ij}|$ and a Jarlskog phase --- three mixing angles and one CP phase --- and whose non-triviality
    $[\Upi] \neq [I]$ is controlled exactly by the transverse component of the metaplectic generator.
    On the derived real cascade that transverse component is absent, so $[\Upi] = [I]$ and no physical mixing
    is produced at this stratum; the only route left open to a non-trivial class is a genuine complex
    non-central metaplectic phase, not derived here, observable only relative to a second fermionic sector.
    All operator identities are verified by exact symbolic computation, with no sampling.
  \end{abstract}

  \noindent\textbf{Keywords:}
  projected Yukawa operator; hermitian square; polar decomposition; chiral polar factor; polar class;
  rephasing invariants; Jarlskog phase; generation mixing; generation levels; partial no-go; Schur residue;
  CP-real branch; non-injective projection.

  \newpage

  \tableofcontents

  \newpage

  \section{Position and purpose}
    \label{sec:position}

    The line note~\cite{Beau2026pyl} closed the first stage of the projected Yukawa sector.
    It identified the determinant line $\LY = \wedge^{2}(\Sbundle)$ as the unique functorial coupling line,
    invisible to the $\SU(2)_{L}$ adjoint sector and carrying the abelian weight, and it read the generation
    levels off the squared projective residue $\Epi^{2}|_{\Cgen} = \diag(1, \tfrac12 + u, \tfrac12 - u)$.
    It also stated, but did not analyse, that the mass matrix is a distinct object: the projected Yukawa
    morphism $\Ypi : \SL \otimes \LY \to \SR$, whose norm, the sign of $u$, and the mixing phase remained
    open.

    This note analyses that morphism, and the analysis begins with a negative result.
    The squared projective residue is a hermitian object on a single chirality, while $\Ypi$ ties the two
    chiralities; the chiral connecting datum is therefore not visible to $\Epi^{2}$.
    Concretely, $\Epi^{2}|_{\Cgen}$ determines the hermitian square $\Hpi = \Ypi^{\dagger}\Ypi$, hence the
    squared Yukawa levels, but leaves $\Ypi$ free up to a unitary chiral polar factor.
    The purpose of the note is to make this precise, to extract exactly what is nonetheless fixed --- the
    spectrum of $\Ypi^{\dagger}\Ypi$ --- and to delimit what is not.
    The discipline of the projective-residue reduction is kept throughout: $\Epi = -\Pi_{S} D (1 - P) D
    \Pi_{S}^{*} = -M^{\dagger} M$ is negative semi-definite, with the split $u$ carried by the transported
    chiral defect~\cite{Beau2026prs}, so a semi-definite operator must not be promoted by hand to an
    oriented chiral morphism.

  \section{The Yukawa operator as a chiral morphism}
    \label{sec:morphism}

    The projected Yukawa coupling is a morphism of the admissible spinor sectors twisted by the coupling
    line,
    \begin{equation}
      \label{eq:yukawa-morphism}
      \Ypi : \SL \otimes \LY \longrightarrow \SR ,
    \end{equation}
    from the left chiral generation carrier, twisted by the determinant line $\LY$, to the right chiral
    carrier.
    The mass matrix is the value of $\Ypi$ on the locked configuration, and the two physically invariant
    hermitian objects are the left and right squares $\Ypi^{\dagger}\Ypi$ on $\SL \otimes \LY$ and
    $\Ypi\,\Ypi^{\dagger}$ on $\SR$.
    The squared projective residue $\Epi^{2}$ is a hermitian operator on a single chirality; it can coincide
    with one of these squares, but it cannot by itself carry the chiral connecting map between them.
    This is the structural origin of the polar ambiguity below.

  \section{The squared Yukawa observable}
    \label{sec:squared}

    \begin{proposition}[Squared Yukawa observable]
      \label{prop:squared-yukawa}
      Let $\Ypi : \SL \otimes \LY \to \SR$ be the projected Yukawa morphism.
      Then the generation-level observable determined by the Schur-residue sector is the positive hermitian
      operator
      \[
        \Hpi := \Ypi^{\dagger}\Ypi
      \]
      on the left chiral generation carrier.
      On the gauge-singlet triplet $\Cgen$, its dimensionless spectral part is fixed, up to the overall
      Yukawa norm $\lamY$, by
      \[
        \Hpi\big|_{\Cgen} = \lamY^{2}\,\diag\!\left(1,\ \tfrac12 + u,\ \tfrac12 - u\right) ,
      \]
      so that $\Spec\big(\Ypi^{\dagger}\Ypi\big) = \lamY^{2}\,\{1,\ \tfrac12 + u,\ \tfrac12 - u\}$.
      Thus $\Epi^{2}|_{\Cgen}$ determines the squared Yukawa levels, not the Yukawa morphism itself.
    \end{proposition}

    \begin{proof}
      The Schur reduction gives $\Epi = -M^{\dagger} M$ negative semi-definite, so $\Epi^{2}$ is a positive
      hermitian operator and equals the square of $-\Epi$; on the triplet, $\Epi^{2}|_{\Cgen} = \diag(1,
      \tfrac12 + u, \tfrac12 - u)$ by the normal form of~\cite{Beau2026prs} with the even sector closed by
      Born--Infeld parity~\cite{Beau2026a34}, and the level reading of~\cite{Beau2026pyl}.
      Identifying the Schur-residue generation observable with the left Yukawa square, $\Hpi = \Ypi^{\dagger}
      \Ypi = \lamY^{2}\,(-\Epi/|\Epi|)$-normalised, gives $\Hpi|_{\Cgen} = \lamY^{2}\,\diag(1, \tfrac12 + u,
      \tfrac12 - u)$ with $\lamY$ the undetermined overall norm.
      The eigenvalues are the diagonal entries, so the spectrum is as stated, and a hermitian square fixes a
      hermitian operator, not the underlying morphism, which is the content of Section~\ref{sec:polar}.
    \end{proof}

    The positive square root is the projective residue itself up to sign: since $\Epi|_{\Cgen}$ is negative
    semi-definite, $\big(-\Epi|_{\Cgen}\big)^{2} = \Epi^{2}|_{\Cgen}$, so $-\Epi|_{\Cgen}$ is the positive
    root $\Hpi^{1/2}/\lamY$.
    No oriented chiral information has been used, in keeping with the reduction discipline.

  \section{Polar decomposition and the undetermined chiral phase}
    \label{sec:polar}

    \begin{proposition}[Chiral polar ambiguity]
      \label{prop:polar-ambiguity}
      The operator $\Hpi = \Ypi^{\dagger}\Ypi$ determines $\Ypi$ only up to a polar factor:
      \[
        \Ypi = \Upi\,\Hpi^{1/2} ,
      \]
      where
      \[
        \Upi : \SL \otimes \LY \longrightarrow \SR
      \]
      is a unitary chiral polar map on the support of $\Hpi$.
      Consequently the squared residue fixes the generation levels but not the complex mixing phase, nor the
      full Yukawa operator.
    \end{proposition}

    \begin{proof}
      For a positive hermitian $\Hpi$ with $\Hpi^{1/2}$ its positive square root, any $\Ypi$ with
      $\Ypi^{\dagger}\Ypi = \Hpi$ satisfies $\Ypi = \Upi\,\Hpi^{1/2}$ with $\Upi := \Ypi\,\Hpi^{-1/2}$ a
      partial isometry, unitary on the support of $\Hpi$, since $\Upi^{\dagger}\Upi = \Hpi^{-1/2}
      \Ypi^{\dagger}\Ypi\,\Hpi^{-1/2} = \Hpi^{-1/2}\Hpi\,\Hpi^{-1/2} = \mathbf{1}$ on that support.
      Conversely, for any such $\Upi$ the product $\Upi\,\Hpi^{1/2}$ has square $\Hpi^{1/2}\Upi^{\dagger}
      \Upi\,\Hpi^{1/2} = \Hpi$, so the solution set is exactly $\{\Upi\,\Hpi^{1/2}\}$.
      The map $\Upi$ acts on the right carrier $\SR$, and $\Hpi$ is invariant under $\Ypi \mapsto V\,\Ypi$
      for every unitary $V$ on $\SR$, $(V\Ypi)^{\dagger}(V\Ypi) = \Ypi^{\dagger} V^{\dagger} V\,\Ypi =
      \Ypi^{\dagger}\Ypi$, so distinct $\Upi$ give distinct $\Ypi$ with identical $\Hpi$.
      Hence $\Hpi$, equivalently $\Epi^{2}|_{\Cgen}$, does not determine $\Ypi$: the generation-mixing
      content and the relative phases reside in $\Upi$, on which $\Epi^{2}$ has no hold.
    \end{proof}

    \begin{remark}[The partial no-go, sharply]
      \label{rem:no-go}
      Proposition~\ref{prop:polar-ambiguity} is a partial no-go: $\Epi^{2}$ does not fix $\Ypi$.
      What survives is exactly the spectral datum of Proposition~\ref{prop:squared-yukawa},
      $\Spec(\Ypi^{\dagger}\Ypi) = \lamY^{2}\{1, \tfrac12 + u, \tfrac12 - u\}$.
      The undetermined factor is not a numerical residue but a unitary chiral map, the carrier of the chiral
      orientation, the relative generation phases, and the mixing.
      The positive scale $\lamY$ is a separate normalisation of the hermitian square $\Hpi$, invisible to the
      dimensionless level operator $\Epi^{2}|_{\Cgen}$, and it is not carried by the unitary $\Upi$.
    \end{remark}

    \begin{remark}[The open datum in $u$ is the orientation branch, not the scalar sign]
      \label{rem:orientation}
      In an oriented $\Jthree$-labelled generation basis $(e_0, e_+, e_-)$ the level operator
      $\diag(1, \tfrac12 + u, \tfrac12 - u)$ already sees the relative sign of $u$.
      The operator $\Hpi$ therefore fixes the sign of $u$ only after such an oriented basis has been chosen.
      Without that chiral-orientation choice the exchange $e_+ \leftrightarrow e_-$ sends $u \mapsto -u$ and
      leaves the unordered squared spectrum unchanged.
      Thus the remaining datum is the physical orientation branch of the split, not an additional scalar
      visible to $\Epi^{2}$ alone.
    \end{remark}

  \section{Diagonal CP-real branch}
    \label{sec:cp-real}

    The reduction~\cite{Beau2026prs} imposes $v = 0$ on the CP-even real sector, the off-diagonal mixing
    $R_{\mathrm{mix}}$ vanishing for phase-free data, with a non-zero $v$ requiring a complex metaplectic
    phase.
    In the present language this is the branch $\Upi = \mathbf{1}$.
    There the Yukawa operator is the positive root itself,
    \begin{equation}
      \label{eq:cp-real}
      \Ypi^{\mathrm{CP}} = \Hpi^{1/2} = \lamY\,\diag\!\left(1,\ \sqrt{\tfrac12 + u},\ \sqrt{\tfrac12 - u}\right),
    \end{equation}
    real, diagonal in the generation basis, with no mixing.
    This is a choice of branch, not a derived fact: a non-trivial $\Upi$ --- a generation rotation or a
    relative phase --- leaves $\Hpi$ and the levels untouched while injecting off-diagonal and complex
    structure that $\Epi^{2}$ cannot register.
    The mixing and the CP phase therefore live entirely in $\Upi$, and the only route to them left open here is
    the complex metaplectic phase of the Lorentzian completion, the same datum that governs the chiral-frontier
    normalisation~\cite{Beau2026aar} and the residual orientation~\cite{Beau2026bim}; none of it is supplied
    here.

  \section{Polar class of the projected Yukawa operator}
    \label{sec:polar-class}

    The polar factor $\Upi$ of Section~\ref{sec:polar} is not a canonical observable before one quotients by
    the admissible chiral rephasings.
    The generation-level operator $\Hpi|_{\Cgen} = \lamY^{2}\,\diag(1, \tfrac12 + u, \tfrac12 - u)$ has a
    distinct spectrum for $0 < u < \tfrac12$, and by the antiunitary $\JPi$ exchange of chiralities the right
    carrier carries the level operator $\Ypi\,\Ypi^{\dagger} = \Upi\,\Hpi\,\Upi^{\dagger}$ with the same
    spectrum~\cite{Beau2026prs}.
    Distinctness pins each carrier to its $\Jthree$-labelled level eigenbasis up to a diagonal phase, so the
    residual admissible freedom is the diagonal rephasing of the generation states.

    \begin{proposition}[Polar class]
      \label{prop:polar-class}
      Let $V_{L} \in \UU(1)^{3}$ act on $\SL \otimes \LY$ and $V_{R} \in \UU(1)^{3}$ on $\SR$ as diagonal
      rephasings of the level eigenbases.
      Then $\Ypi \mapsto V_{R}\,\Ypi\,V_{L}^{-1}$, $\Hpi \mapsto V_{L}\,\Hpi\,V_{L}^{-1}$, and
      \[
        \Upi \longmapsto V_{R}\,\Upi\,V_{L}^{-1} ,
      \]
      so the projected Yukawa construction carries not a preferred matrix $\Upi$ but its rephasing class, the
      double quotient
      \[
        [\Upi] \in \UU(1)^{3}_{R} \,\backslash\, \UU(3) \,/\, \UU(1)^{3}_{L} .
      \]
      Its rephasing invariants are the moduli $|(\Upi)_{ij}|$ together with the Jarlskog-type phase
      \[
        J_{\Pi} = \Im\!\big( (\Upi)_{11}(\Upi)_{22}\,\overline{(\Upi)_{12}}\,\overline{(\Upi)_{21}} \big) .
      \]
      The common diagonal phase acts trivially on the double action, so the dimension count is $9 - (2\cdot 3 -
      1) = 4$, namely three mixing angles and one CP phase.
    \end{proposition}

    \begin{proof}
      If the right carrier admitted an arbitrary $\UU(3)$ rotation, the choice $V_{R} = \Upi$ would give
      $V_{R}^{-1}\,\Ypi = \Hpi^{1/2}$, diagonal and positive, removing $\Upi$ entirely and leaving only
      $\Spec(\Hpi)$.
      The distinct right level operator $\Ypi\,\Ypi^{\dagger}$ forbids this: a unitary that commutes with a
      diagonal of distinct entries is diagonal, since $(WD - DW)_{ij} = W_{ij}(d_{j} - d_{i})$, so the
      admissible $V_{L}, V_{R}$ are diagonal rephasings only.
      Under such rephasings $|(V_{R}\,\Upi\,V_{L}^{-1})_{ij}| = |(\Upi)_{ij}|$ and the quartet phase $J_{\Pi}$
      are invariant, and for $n = 3$ these exhaust the invariants; the dimension of $\UU(3)$ is nine, the
      rephasings remove $2\cdot 3 - 1 = 5$, and four physical parameters survive, three angles and one phase.
    \end{proof}

    The next obstruction is therefore not the existence of $\Upi$, but the non-triviality of its class: whether
    the complex metaplectic phase produces $[\Upi] \neq [I]$ or only a representative diagonal-rephasing
    equivalent to the identity.

    \begin{proposition}[Polar non-triviality]
      \label{prop:polar-nontriviality}
      Write $\Upi(\gamma) = \exp(\gamma A)$ for the chiral generator $A = \partial_{\gamma}\Upi|_{0}$ induced by
      the complex metaplectic phase $\gamma$, with $A$ anti-hermitian.
      The diagonal-rephasing orbit through the identity has tangent space the diagonal anti-hermitian
      matrices, so $[\Upi(\gamma)] \neq [I]$ at first order if and only if $A$ has a non-zero off-diagonal
      (transverse) part.
      The rephasing-invariant detectors may be taken as
      \[
        |(\Upi)_{ij}|^{2} = \gamma^{2}|A_{ij}|^{2} + O(\gamma^{3}), \qquad i \neq j ,
      \]
      or, equivalently for a positive branch of $\gamma$, $\partial_{\gamma}^{+}|(\Upi)_{ij}|_{\gamma = 0} =
      |A_{ij}|$, so the off-diagonal moduli detect precisely the transverse component of the metaplectic
      generator; the Jarlskog invariant has leading non-vanishing order $\gamma^{3}$ and is proportional to
      $\Im(A_{12}A_{23}A_{31})$, independent of the diagonal phases.
    \end{proposition}

    \begin{proof}
      A curve $V_{R}(s)\,V_{L}(s)^{-1}$ through the identity in the rephasing group has derivative
      $i(D_{R} - D_{L})$ at $s = 0$, diagonal and anti-hermitian, and every diagonal anti-hermitian is so
      reached; hence the orbit tangent at $I$ is exactly the diagonal anti-hermitian subspace.
      For $\Upi(\gamma) = \exp(\gamma A)$ the off-diagonal entry is $(\Upi)_{ij} = \gamma A_{ij} + O(\gamma^{2})$
      so $|(\Upi)_{ij}|^{2} = \gamma^{2}|A_{ij}|^{2} + O(\gamma^{3})$, and the moduli detect precisely the
      transverse part of $A$.
      The quartet phase vanishes at orders $\gamma^{0}, \gamma^{1}, \gamma^{2}$ and its leading $\gamma^{3}$
      coefficient is proportional to $\Im(A_{12}A_{23}A_{31})$, the genuine three-generation datum, with no
      dependence on the diagonal phases.
    \end{proof}

    \begin{corollary}[Real-cascade collapse]
      \label{cor:real-cascade-collapse}
      On the derived real cascade the off-diagonal mixing channel is absent, $v = 0$, the transported chiral
      defect being the real $V\!-\!A$ diagonal split~\cite{Beau2026prs, Beau2026ebj}; hence $A$ is diagonal,
      $\Upi(\gamma)$ stays diagonal, and $[\Upi] = [I]$, with $|(\Upi)_{ij}| = \delta_{ij}$ and $J_{\Pi} = 0$.
      No physical mixing is produced at this stratum.
      A non-trivial class $[\Upi] \neq [I]$, hence CKM/PMNS-type mixing, requires a transverse metaplectic
      component.
      In the present corpus-derived real cascade this transverse component is absent; the only route left open
      is therefore a genuine complex metaplectic phase, not derived here and still conditional on an
      unprescribed phase and ordering~\cite{Beau2026bim}.
      The resulting mixing matrix is physical only as a relative polar class between two fermionic sectors
      sharing a generation carrier.
    \end{corollary}

    \begin{remark}[Two transverse blocks]
      \label{rem:two-transverse-blocks}
      The transverse obstruction has two distinct components.
      On the generation triplet $\Cgen = \langle e_{0}, e_{+}, e_{-}\rangle$ the polar tangent transverse to the
      diagonal-rephasing orbit decomposes into the internal block $e_{0} \leftrightarrow e_{\pm}$ and the
      external block $e_{+} \leftrightarrow e_{-}$, the latter denoted $R_{\mathrm{mix}}$.
      For a single metaplectic $\mathfrak{sl}_{2}$ cascade lifted to $\Sym^{2}$, the correct unitary chiral
      generator is the anti-hermitian part of the $\JPi$-odd component,
      \[
        A_{\Pi} = \Pi_{\mathrm{antiherm}}\,\Pi_{\JPi\text{-odd}}\big(\dot{\Upi}\big) .
      \]
      On real cascade data this generator vanishes, recovering Corollary~\ref{cor:real-cascade-collapse}.
      If the $\mathfrak{sl}_{2}$ coefficients acquire a genuinely complex non-central part, the internal block
      $e_{0} \leftrightarrow e_{\pm}$ is sourced.
      By contrast the external block $e_{+} \leftrightarrow e_{-}$ is never sourced by the
      $\Sym^{2}(\mathfrak{sl}_{2})$ lift alone; it lies outside the single-cascade metaplectic image and would
      require an additional tower datum.
      The chiral-frontier normalisation channel and the polar-mixing channel therefore separate: the oriented
      area $N_{A}$ belongs to the real Cartan / $\Jthree$ channel~\cite{Beau2026aar} and can normalise the
      diagonal split without predicting a CKM/PMNS polar class.
      A non-trivial polar class in the present carrier first requires a complex non-central metaplectic
      component, and the external $R_{\mathrm{mix}}$ block requires still more: a generator beyond the single
      $\mathfrak{sl}_{2}$ metaplectic cascade.
    \end{remark}

    \begin{proposition}[Reality structure of the polar generator]
      \label{prop:reality-structure}
      Write the non-central metaplectic step as $M = p\,E + q\,F + r\,H$ in the real $\mathfrak{sl}_{2}$ basis,
      and let $\sigma(M) = \overline{M}$ be the entrywise conjugation whose fixed real form is
      $\mathfrak{sl}_{2}(\mathbb{R}) = \mathrm{span}_{\mathbb{R}}\{E, F, H\}$, so $\sigma(M) = M$ iff
      $p, q, r \in \mathbb{R}$.
      The chiral polar generator $A_{\Pi} = \Pi_{\mathrm{antiherm}}\Pi_{\JPi\text{-odd}}(\dot{\Upi})$ depends only
      on the $\sigma$-anti-real part $\tfrac12(M - \sigma M) = i\,(\Im p\,E + \Im q\,F + \Im r\,H)$, and splits as
      the polar class does.
      Its transverse (off-diagonal) part is non-zero iff $\Im p \neq 0$ or $\Im q \neq 0$, while its diagonal part
      is the pure $\Jthree$ rephasing $i\,\Im r\,\diag(0, 2, -2)$, which lies in the $\UU(1)^{3}$ double-coset
      quotient defining $[\Upi]$ and is therefore class-trivial.
      Hence
      \[
        [\Upi] = [I] \quad\Longleftrightarrow\quad \Im p = \Im q = 0 ,
      \]
      independently of $\Im r$.
      In particular a complex representation matrix obtained from a complex spinorial frame $W$ leaves the class
      unchanged, $A_{\Pi} \mapsto W A_{\Pi} W^{\dagger}$, so the mixing is controlled by the reality of the
      \emph{coefficients} $p, q$, not by that of the matrix entries.
    \end{proposition}

    \begin{corollary}[Negative closure at the present stratum]
      \label{cor:negative-closure-present-stratum}
      At the present full-tower stratum the non-central step has real coefficients.
      The cascade step is $g = \exp(tE)\exp(sF)$ with $t, s \in \mathbb{R}$, its per-step $\JPi$-odd $\Jthree$
      component being the real oriented area $\alpha = ts$~\cite{Beau2026prs, Beau2026aar}; and the inherited lift
      carries only the central cyclotomic phase $\zeta_{q}^{\Delta A_{c}}$, with no Fourier or Weyl generator and
      hence no metaplectic Gauss phase~\cite{Beau2026cho}, a central datum that is projectively trivial for the
      polar factor.
      Thus $\Im p = \Im q = 0$ is forced and, by Proposition~\ref{prop:reality-structure}, $A_{\Pi} = 0$ and
      $[\Upi] = [I]$.
      The diagonal split nevertheless stands: $N_{A} \neq 0$~\cite{Beau2026cho} and $u \neq 0$~\cite{Beau2026prs}.
      The present stratum therefore derives the diagonal generation split but not the mixing.
      A non-trivial internal block requires a genuinely complex non-central metaplectic phase, a Weil generator
      absent from the present pipeline and available only in a new tower stratum, gated by the same full-tower
      type-rigidity~\cite{Beau2026aog} as the capacity normalisation.
    \end{corollary}

    \begin{remark}[Two delimitations]
      \label{rem:precautions}
      First, the spin-stratum type-rigidity of~\cite{Beau2026aog} closes the orientation caveat
      $[H\text{-orient}]$ and excludes a spin-Galois $\sqrt{5}$ factor orthogonal to the central phase, but it
      does not bear on the non-triviality of $[\Upi]$: chiral orientation and polar class are distinct data.
      Second, the generation split $u \neq 0$~\cite{Beau2026prs, Beau2026bim} is a diagonal-channel datum and
      produces no mixing; it contributes nothing to the off-diagonal moduli or to $J_{\Pi}$.
      Mixing is the transverse polar datum, not the diagonal split.
    \end{remark}

  \section{Status of results}
    \label{sec:status}

    \paragraph{Derived (this note).}
      The generation-level observable of the Schur-residue sector is the positive hermitian square $\Hpi =
      \Ypi^{\dagger}\Ypi$, with $\Spec(\Hpi)|_{\Cgen} = \lamY^{2}\{1, \tfrac12 + u, \tfrac12 - u\}$
      (Proposition~\ref{prop:squared-yukawa}).
      The Yukawa morphism is determined by this square only up to a unitary chiral polar factor, $\Ypi =
      \Upi\,\Hpi^{1/2}$ (Proposition~\ref{prop:polar-ambiguity}).
      The polar factor is moreover not canonical but a rephasing class, the double quotient $[\Upi] \in
      \UU(1)^{3}_{R} \backslash \UU(3) / \UU(1)^{3}_{L}$, with invariants the moduli $|(\Upi)_{ij}|$ and the
      Jarlskog phase $J_{\Pi}$, four physical parameters in all (Proposition~\ref{prop:polar-class}).
      Its non-triviality is controlled by the transverse component of the metaplectic generator: $[\Upi] \neq
      [I]$ iff that generator has a non-zero off-diagonal component (Proposition~\ref{prop:polar-nontriviality}).
      All of these are structural and unconditional given the carrier identifications of~\cite{Beau2026prs,
      Beau2026a34, Beau2026pyl}.

    \paragraph{Conditional.}
      On the derived real cascade the off-diagonal channel is absent, $v = 0$, so $[\Upi] = [I]$ and no
      physical mixing is produced at this stratum (Corollary~\ref{cor:real-cascade-collapse}).
      The transverse datum further splits into two blocks (Remark~\ref{rem:two-transverse-blocks}): the internal
      block $e_{0} \leftrightarrow e_{\pm}$ is sourcable by a genuinely complex non-central
      $\mathfrak{sl}_{2}$ metaplectic coefficient, while the external block $e_{+} \leftrightarrow e_{-}$ is not
      in the image of the single-cascade $\Sym^{2}(\mathfrak{sl}_{2})$ lift.
      Consequently $N_{A}$ may fix the diagonal oriented-area normalisation without fixing the polar mixing
      class.
      At the present full-tower stratum these non-central coefficients are moreover forced real --- the cascade
      step is real~\cite{Beau2026prs, Beau2026aar} and the inherited lift carries only the central
      phase~\cite{Beau2026cho} --- so $\Im p = \Im q = 0$, $A_{\Pi} = 0$ and $[\Upi] = [I]$: the stratum derives
      the diagonal split but not the mixing (Proposition~\ref{prop:reality-structure},
      Corollary~\ref{cor:negative-closure-present-stratum}).

    \paragraph{Open.}
      The positive Yukawa norm $\lamY$ and a non-trivial polar class $[\Upi] \neq [I]$ are not fixed here.
      The norm $\lamY$ scales $\Hpi$, whereas a non-trivial $[\Upi]$ requires a genuinely complex non-central
      metaplectic phase, absent from the present pipeline and available only through a new tower
      stratum~\cite{Beau2026aog}.
      It is observable as a mixing matrix only relative to a second fermionic sector.
      Hence the absolute masses, the generation mixing, and the CP structure remain downstream open data, in
      agreement with~\cite{Beau2026fmnote}.

    \paragraph{Arithmetic gate for the diagonal split.}
      The diagonal split parameter \(u\) and the angular--radial datum \(\varepsilon\) are not
      independent invariants.
      They are the same functional of the level ratio
      \[
        R_\Pi
        =
        \frac{\frac12+u}{\frac12-u}.
      \]
      Indeed, the inverse parametrisation
      \[
        \varepsilon(R)
        =
        \frac{R-1}{2(R+1)}
      \]
      gives \(\varepsilon(R_\Pi)=u\) identically.
      This identity is only a functional identification.
      The representation \(\Sym^{2}(\mathbb{C}^{2})\) does not select the value of \(u\):
      the diagonal split
      \[
        \diag\left(1,\frac12+u,\frac12-u\right)
      \]
      commutes with \(\Jthree\) for every \(u\).
      Thus the polar spectrum of \(\Hpi\) carries the vertical diagonal split as a free
      carrier-preserving datum, but does not by itself carry the horizontal ADE arithmetic
      obstruction.

      The value \(u=\varepsilon=1/10\) is therefore equivalent to the dictionary gate
      \[
        R_\Pi = R_{\mathrm{ADE}}=\frac32,
      \]
      where \(R_{\mathrm{ADE}}\) is the order-\(5\) Cayley-spectral ratio~\cite{Beau2026aog}.
      This gate is compatible with the closed angular datum, but it is not produced by the
      carrier-preserving recursion.
      In particular, the recursion may fix the vertical parity type and the field
      \(\mathbb{Q}(\sqrt5)\) in which the obstruction is visible, but it does not generate the
      horizontal Galois transport \(2\leftrightarrow 2'\) selecting the order-\(5\) value.
      Consequently the notation \(u\sim_{\mathrm{dict}}\varepsilon\) records the current
      status: same functional, distinct carriers, and value-level equality only after the
      ADE dictionary identification.

    \paragraph{Direct level assignment no-go.}
      Assume the arithmetic dictionary gate \(R_\Pi=R_{\mathrm{ADE}}=3/2\).
      Then \(u=1/10\), and the squared Yukawa levels fixed by \(\Hpi\) are
      \[
        \lamY^{2}\left\{1,\frac35,\frac25\right\}.
      \]
      Their dimensionless squared ratios are
      \[
        \frac52:\frac32:1,
      \]
      and the corresponding root-level ratios are
      \[
        1:\sqrt{\frac35}:\sqrt{\frac25}.
      \]
      These ratios are of order one.
      They therefore cannot be identified with the observed charged-fermion mass hierarchy by a direct generation
      assignment: the smallest observed inter-generation step already exceeds the whole order-one spread by two orders
      of magnitude.
      The squared projective residue fixes an internal level split, not the physical mass hierarchy.

    \paragraph{Separation of level split and hierarchy.}
      The diagonal datum \(u\) should consequently be read as a generation-level weight inside the positive hermitian
      square \(\Hpi=\Ypi^{\dagger}\Ypi\), not as a direct mass-ratio predictor.
      The observed hierarchy must be carried by generation-dependent stabilisation depths in the projective
      amplification law, or by a later operator stratum not analysed here.
      It is not carried by a sector-dependent normalisation of \(\lamY\): the Yukawa norm \(\lamY\) remains a single
      downstream scale throughout this note.
      At this stratum, the dictionary gate can fix the order-one split
      \[
        \left\{1,\frac35,\frac25\right\},
      \]
      but it does not fix the charged-fermion hierarchy.

    \paragraph{Hierarchy not carried by the cascade rate.}
      The single-level projective amplification factor is itself of order one: the Kesten--McKay coefficients and the
      ADE representation dimensions give inter-level separations of only a fraction of a log-decade, with spectral
      ratios in \([3/2, 2]\)~\cite{Beau2026a5}.
      Stacked with the order-one split \(\{1, 3/5, 2/5\}\) it therefore stays of order one and cannot reach the
      observed charged-fermion hierarchy.
      The hierarchy is therefore not carried by the order-one split \(u\), nor by the single-level cascade factor
      alone.
      It is carried, if at all, by generation-dependent cascade depths \(n_g\) in the projective amplification law
      \(\Lambda_{\mathrm{proj}}(n)\) in the exponential growth regime~\cite{Beau2026a5}.
      The exponent fixes the amplification rate; the generation stabilisation depths remain the next structural datum.

    \paragraph{Two lemmas: separation and level-crossing cancellation.}
      We record the two structural statements that freeze the present front, both verified by exact symbolic
      computation with no fit (see Section~\ref{sec:repro}).
      Write the mass-channel factorisation in the form
      \begin{equation}
        \label{eq:mass-factorisation}
        m_g \;=\; \lamY\, w_g\, S_g\, A(n_g),
        \qquad A(n_g) = \exp(\beta^{\ast} n_g),
      \end{equation}
      where \(g\in\{1,2,3\}\) is the generation label, \(w_g\) is the Schur-residue weight of the squared
      residue, \(S_g\) is the stratigraphic order-one factor, and \(A(n_g)\) is the depth amplification
      generated along the projective cascade at the exponent \(\beta^{\ast}=1/(\delta+\tfrac12)\)~\cite{Beau2026a20}.

      \begin{lemma}[Separation of the generation label and the Schur-residue weight]
        \label{lem:separation-label-weight}
        On the common rank-three carrier \(\Cgen \simeq \mathrm{Sym}^{2}(V_\rho)\) the generation label and the
        Schur-residue weight are distinct, multiplicatively acting factors.
        The label \(g\in\{1,2,3\}\) is fixed by the three admissible stratigraphic levels of the binary
        icosahedral group~\cite{Beau2026a4}.
        The weights \(w_g\in\{1, \tfrac35, \tfrac25\}\) are the \(J_3\)-triplet of \(\Hpi\) at the dictionary
        gate and are not identified with the stratigraphic level ratios.
      \end{lemma}
      \begin{proof}
        The two factors depend on disjoint data.
        The weight \(w_g\) is a function of the single vertical datum \(u\) alone: the diagonal
        \(\diag(1,\tfrac12+u,\tfrac12-u)\) commutes with \(J_3\) for every \(u\), so the squared-residue split is
        free under \(\mathrm{Sym}^{2}\).
        The stratigraphic factor \(S_g\) is the Kesten--McKay and representation datum \(c_i/\dim\rho_i\) of the
        \(2I\) Cayley spectrum~\cite{Beau2026a5}.
        Their dimensionless ratio-sets differ: \(w_g\) gives \(\{1, \tfrac32, \tfrac52\}\) while the stratigraphic
        levels \(\{20,24,30\}\) give \(\{1, \tfrac65, \tfrac32\}\).
        Hence the two readings cannot be collapsed into a single observable.
      \end{proof}

      \begin{lemma}[Level-crossing depths cannot carry the hierarchy]
        \label{lem:level-crossing-cancellation}
        In the exponential trajectory-branching regime \(\Lambda_{\mathrm{proj}}(n)\simeq\lambda_2(n) =
        L_0\exp(\beta^{\ast} n)\), suppose the generation depths are the level-crossing depths
        \(n_{\mathrm{proj}}(\lambda) = (1/\beta^{\ast})\log(\lambda/L_0)\) at which each stratigraphic level
        \(\lambda_g\) becomes projectable.
        Then the depth amplification of~\eqref{eq:mass-factorisation} satisfies
        \begin{equation}
          \label{eq:level-crossing-cancellation}
          \frac{A(n_i)}{A(n_j)} \;=\; \exp\!\bigl(\beta^{\ast}(n_i-n_j)\bigr) \;=\; \frac{\lambda_i}{\lambda_j},
        \end{equation}
        independently of \(\beta^{\ast}\).
        On the admissible levels \(\{20,24,30\}\) the amplified ratios are therefore bounded by \(3/2\), and the
        level-crossing depths cannot carry the charged-fermion hierarchy.
      \end{lemma}
      \begin{proof}
        Substituting \(n_{\mathrm{proj}}(\lambda) = (1/\beta^{\ast})\log(\lambda/L_0)\) into
        \(A(n)=\exp(\beta^{\ast} n)\) gives \(A(n_{\mathrm{proj}}(\lambda)) = \lambda/L_0\); the exponent cancels
        and the ratio reduces to \(\lambda_i/\lambda_j\).
        For the three admissible levels this is at most \(30/20=3/2\), an order-one quantity.
      \end{proof}

      \begin{corollary}[Intrinsic branching depths]
        \label{cor:intrinsic-branching-depths}
        A charged-fermion hierarchy in~\eqref{eq:mass-factorisation} requires the depths \(n_g\) to be intrinsic
        branching-stabilisation data of the exponential trajectory, decoupled from the stratigraphic level ratios.
        The required gaps \(\Delta n \sim \log(\text{ratio})/\beta^{\ast}\) (of order \(64\) for the \(\tau\!:\!e\)
        step at \(\beta^{\ast}\simeq 0.127\)) are not reachable from the static data
        \(\{\lambda_g,\dim\rho_g,\beta^{\ast}\}\), whose logarithms are of order one.
      \end{corollary}

      \begin{remark}[Scope]
        \label{rem:depths-scope}
        Lemma~\ref{lem:level-crossing-cancellation} does not derive \(n_g\); it proves where \(n_g\) cannot come
        from.
        The hierarchy is thereby reduced exactly to the depth-triplet problem for \(\lambda_2(n)\) along the
        admissible cascade, with the depths isolated as the carrier.
        Deriving an intrinsic branching-stabilisation law from the growth law \(N(\lambda;n)\) of~\cite{Beau2026a5}
        is a separate, later front.
      \end{remark}

    \paragraph{Summary.}
      Front~3b is closed as a squared-Yukawa-levels result, not a mixing result, and Front~3c characterises
      the chiral polar factor as a rephasing class whose non-triviality is the transverse metaplectic datum.
      The squared residue $\Epi^{2}|_{\Cgen}$ can close $\Hpi = \Ypi^{\dagger}\Ypi$; it cannot close $\Ypi$
      unless the class $[\Upi]$ is fixed, and that class collapses to the identity on the derived real cascade.

  \section{Reproducibility}
    \label{sec:repro}

    All operator identities are verified by exact symbolic computation, with no floating-point sampling.
    The squared-observable results of Sections~\ref{sec:squared}--\ref{sec:cp-real} ---
    the level identification and spectrum of Proposition~\ref{prop:squared-yukawa}, the PRS sign consistency
    $(-\Epi|_{\Cgen})^{2} = \Epi^{2}|_{\Cgen}$, the polar non-uniqueness of
    Proposition~\ref{prop:polar-ambiguity} (for a generic positive $\Hpi$ and an arbitrary unitary, both a
    generation rotation and a diagonal phase leave $\Ypi^{\dagger}\Ypi$ invariant while changing $\Ypi$),
    the invariance of $\Hpi$ under right-carrier unitaries, and the diagonal no-mixing CP-real branch ---
    are checked in \texttt{simulation/fermionic-matter/front3b\_yukawa\_operator.py}.
    The polar-class results of Section~\ref{sec:polar-class} ---
    the diagonal-rephasing class transformation and free-right triviality, the distinct-spectrum pinning and
    the diagonal commutant, the invariance of the moduli and of $J_{\Pi}$ and the four-parameter count
    (Proposition~\ref{prop:polar-class}), the orbit tangent at the identity, the modulus and Jarlskog
    detectors of the transverse part (Proposition~\ref{prop:polar-nontriviality}), and the diagonal
    real-cascade collapse (Corollary~\ref{cor:real-cascade-collapse}) ---
    are checked in \texttt{front3c\_polar\_class\_audit.py} and
    \texttt{front3c\_polar\_class\_nontriviality.py}.
    The two-block decomposition of Remark~\ref{rem:two-transverse-blocks} ---
    the vanishing of the chiral generator $A_{\Pi}$ on real cascade data, the sourcing of the internal block
    $e_{0} \leftrightarrow e_{\pm}$ by a complex non-central coefficient, the identical vanishing of the
    external block $R_{\mathrm{mix}}$ in the $\Sym^{2}(\mathfrak{sl}_{2})$ lift, and the independence of the
    mixing datum from the real oriented area $N_{A}$ ---
    is checked in \texttt{front3d\_transverse\_route.py}.
    The reality-structure results of Section~\ref{sec:polar-class} ---
    the dependence of $A_{\Pi}$ on the $\sigma$-anti-real part alone, the equivalence
    $[\Upi] = [I] \Leftrightarrow \Im p = \Im q = 0$, the class-triviality of the imaginary $\Jthree$ rephasing,
    the projective triviality of the central scalar, and the spinorial-frame covariance distinguishing matrix
    complexity from coefficient complexity (Proposition~\ref{prop:reality-structure},
    Corollary~\ref{cor:negative-closure-present-stratum}) ---
    are checked in \texttt{front3e\_reality\_structure.py}.
    The separation and level-crossing results of Section~\ref{sec:status} ---
    the distinctness of the Schur-residue weight and the stratigraphic factor on the common rank-three carrier
    (Lemma~\ref{lem:separation-label-weight}), the \(\beta^{\ast}\)-independent cancellation
    \(A(n_i)/A(n_j)=\lambda_i/\lambda_j\) of the level-crossing depths
    (Lemma~\ref{lem:level-crossing-cancellation}), and the unreachability of the required depth gaps from the
    static data (Corollary~\ref{cor:intrinsic-branching-depths}) ---
    are checked in \texttt{front\_depths\_ng\_reconciliation.py}.

  \newpage

  \bibliographystyle{plain}
  \bibliography{cosmochrony-bibliography}

\end{document}
